R Notebook

### 2021-03-11 eda.Rmd
EDA work on more variables

- Filtered data such that `api_temp = oxymorphone`
- Dropped `country` and `api_temp`
- Removed observations with missing `ppm` (14)
- Dropped unused levels
- Dropped observations where `state == 'USA'`

@@

# Only 1 country (USA), and we're only working with 
drop_cols <- c('country', 'api_temp')
streetrx <- streetrx %>% select(-any_of(drop_cols))
```

@@

@@

>>>>

- Filtered data such that `api_temp = oxymorphone`
- Dropped `country`, `api_temp`, and 'form_temp`
- Removed observations with missing `ppm` (14)
- Dropped unused levels
- Dropped observations where `state == 'USA'`

@@

# 1 country (USA), 1 drug (oxymorphone), and always as a pill/tablet (except once)
drop_cols <- c('country', 'api_temp', 'form_temp')
streetrx <- streetrx %>% select(-any_of(drop_cols))
```

@@

## yq_pdate

This variable contains the year of sale as well as the quarter (1-4). We already have the year extracted, let's get the quarter. First are all the dates as expected?

```{r}
summary(streetrx$yq_pdate)
```

Well, we certainly have a year that is likely wrong (1969), but at least all of these values contain 5 digits

```{r}
# To find the final number (quarter), simply mod 10
streetrx$quarter <- streetrx$yq_pdate %% 10
```

Let's take a look at the density:

```{r}
small_ppm <- streetrx %>% filter(ppm < 10)
ggplot(small_ppm, aes(x=ppm, color=as.factor(quarter))) + geom_density(bw=0.8)
```

## Price Date

Firstly, change this to a date and extract its component parts:

```{r}
streetrx$price_date <- as.Date(streetrx$price_date, format= '%m/%d/%y')
streetrx <- streetrx %>% mutate(year = lubridate::year(price_date),
                                month = lubridate::month(price_date),
                                day = lubridate::day(price_date))
```

@@

## mgstr

Dosage strength in mg of the units purchased. Let's check a summary:

```{r}
summary(streetrx$mgstr)
```

Not actually a continuous variable:

```{r}
table(streetrx$mgstr)
```

## bulk_purchase

```{r}
table(streetrx$bulk_purchase)
```

```{r}
ggplot(small_ppm, aes(x=ppm, fill=bulk_purchase, color=bulk_purchase)) +
    geom_histogram(position="identity", alpha=0.5)
```
